What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts

### 2023-10-26 SDSS-2023/testing-charts-accuracy.qmd
figure 8 - horizontal labels

library(readr)
library(tidyverse)
library(knitr)
library(bibtex)
theme_set(theme_bw())
theme_update(axis.ticks = element_blank())

@@

@@

acs_educ5 <- data.frame(educ5 = c("Less than HS", "HS or equivalent", "Some college",  "Bachelor", "Post graduate"), 
                      estimate = c(
                      3636691 + 10793507 + 14256849, 
                      9974372 + 59636386,
                      13050667 + 45042031 + 19614710,
                      3677359 + 46354331,
                      29454503),
                      moe_percent = c(sqrt(0.1^2+0.1^2+0.1^2),
                              sqrt(0.1^2+0.1^2 ),
                              sqrt(0.1^2+0.1^2 + 0.1^2),
                              sqrt(0.1^2+0.2^2 ),
                              0.1) 
                      )
acs_educ5 <- acs_educ5 %>% mutate(percent = estimate/(30339089+225152317)*100)

omni_educ <- data.frame(svyby(~educ5, by=~chart, design=survey_123, svymean))
omni_educ_long <- omni_educ %>% 
  pivot_longer(starts_with("educ"), names_to="educ", values_to = "mean") %>% 
  dplyr::select(chart, educ, mean) 
omni_educ_long$SE <- omni_educ %>% pivot_longer(starts_with("se."), names_to="educ", values_to="SE") %>% pull(SE)
omni_educ <- omni_educ_long %>% mutate(
  educ = gsub("educ5", "", educ),
  educ = gsub(".", " ", educ, fixed=TRUE)
)  
omni_educ <- omni_educ %>% mutate(
  educ = factor(educ, levels = c("Less than HS", "HS or equivalent", "Some college", "Bachelor", "Post graduate"))
)

@@

@@

One contributing factor to the observed patterns in response accuracy might be the way that participants interact with the different designs.
Across all tasks, about half of all participants make use of the option to zoom into charts.
We observe that while zooming does help with the overall accuracy (which is in agreement with the findings by @luModelingJustNoticeable2022 about the physical size of stimuli), the increase is not significant.
However, different designs lead to different rates of zooming: we observe in \autoref{fig-zoom-123} that when dealing with the vertical design, the rate of zooming is significantly higher than for the two horizontal designs.

@@

```{r}
#| label: tbl-zoom-123
#| tbl-cap: Coefficients for logistic regression of zooming by task

@@

coefs_zoom$term <- c("$$\\hat{\\mu}$$", "$$\\text{They are the same } \\widehat{\\rho}_{2}$$", 
                     "Wrong $$\\widehat{\\rho}_{3}$$", "Aligned $$\\widehat{\\tau}_{2}$$", 
                     "Horizontal wide $$\\widehat{\\delta}_{2}$$", "Vertical $$\\widehat{\\delta}_{3}$$")

@@

@@

gender_cd + age_cd  + educ_cd +income_cd + 
    plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4))

gender_ab + age_ab  + educ_ab +income_ab + 
  plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4))
```

>>>>

library(readr)
library(tidyverse)
library(tidycensus)
library(knitr)
library(bibtex)
theme_set(theme_bw())
theme_update(axis.ticks = element_blank())

@@

#levels(round_123$educ5) <- c("< HS", "HS", "Some\ncollege", "Bachelor", "Post\ngraduate")

@@

acs_educ5 <- data.frame(educ5 = levels(round_123$educ5), 
                      estimate = c(
                      3636691 + 10793507 + 14256849, 
                      9974372 + 59636386,
                      13050667 + 45042031 + 19614710,
                      3677359 + 46354331,
                      29454503),
                      moe_percent = c(sqrt(0.1^2+0.1^2+0.1^2),
                              sqrt(0.1^2+0.1^2 ),
                              sqrt(0.1^2+0.1^2 + 0.1^2),
                              sqrt(0.1^2+0.2^2 ),
                              0.1) 
                      )
acs_educ5 <- acs_educ5 %>% mutate(percent = estimate/(30339089+225152317)*100)

omni_educ <- data.frame(svyby(~educ5, by=~chart, design=survey_123, svymean))
omni_educ_long <- omni_educ %>% 
  pivot_longer(starts_with("educ"), names_to="educ", values_to = "mean") %>% 
  dplyr::select(chart, educ, mean) 
omni_educ_long$SE <- omni_educ %>% pivot_longer(starts_with("se."), names_to="educ", values_to="SE") %>% pull(SE)
omni_educ <- omni_educ_long %>% mutate(
  educ = gsub("educ5", "", educ),
  educ = gsub(".", " ", educ, fixed=TRUE)
)  
omni_educ <- omni_educ %>% mutate(
  educ = factor(educ, levels = levels(round_123$educ5))
)

@@

levels(acs_income$income4) <- levels(round_123$income4)

@@

One contributing factor to the observed patterns in response accuracy might be the way that participants interact with the different designs.
Across all tasks, about half of all participants make use of the option to zoom into charts.
We observe that while zooming does help with the overall accuracy (which is in agreement with the findings by @luModelingJustNoticeable2022 about the physical size of stimuli), the increase is not significant.
However, different designs lead to different rates of zooming: we observe in \autoref{fig-zoom-123} that when a significantly lower rate of respondents chose to make use of zooming when answering the horizontal designs compared to the vertical design.

@@

```{r results="asis"}
#| label: tbl-zoom-123
#| tbl-cap: Coefficients for logistic regression of zooming by task

@@

# coefs_zoom$term <- c("$$\\hat{\\mu}$$", "$$\\text{They are the same } \\widehat{\\rho}_{2}$$", 
#                      "Wrong $$\\widehat{\\rho}_{3}$$", "Aligned $$\\widehat{\\tau}_{2}$$", 
#                      "Horizontal wide $$\\widehat{\\delta}_{2}$$", "Vertical $$\\widehat{\\delta}_{3}$$")

coefs_zoom$term <- c("\\hat{\\mu}", 
                         "\\text{They are the same } \\widehat{\\rho}_{2}", 
                         "Wrong $$\\widehat{\\rho}_{3}$$", 
                         "Aligned $$\\widehat{\\tau}_{2}$$", 
                         "Horizontal wide $$\\widehat{\\delta}_{2}$$", 
                         "Vertical $$\\widehat{\\delta}_{3}$$")

@@

thm <- 
  theme(axis.text.x = element_text(angle=0, hjust = 0.5))

thm2 <- 
  theme(axis.text.x = element_text(angle=0, hjust = 0.5),
        axis.title.y = element_blank(), 
        axis.text.y = element_blank())

@@

educ_cd <- educ_cd + 
  scale_x_discrete(labels=c("< HS", "HS", "Some\ncollege", "Bachelor", "Post\ngraduate"))
educ_ab <- educ_ab + 
  scale_x_discrete(labels=c("< HS", "HS", "Some\ncollege", "Bachelor", "Post\ngraduate"))

income_cd <- income_cd +
  scale_x_discrete(labels=c("< $30k", "$30-60k", "$60-100k", "≥ $100k"))
income_ab <- income_ab +
  scale_x_discrete(labels=c("< $30k", "$30-60k", "$60-100k", "≥ $100k"))

gender_cd + thm + 
 age_cd  + thm2 + 
  educ_cd  + thm2 + 
  income_cd + thm2 + 
    plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4)) &
  theme(legend.position = "bottom",
        legend.box.margin = margin(t = -15, r = 0, b = 0, l = 0, unit = "pt"))

gender_ab + thm + 
  age_ab  + thm2 + 
  educ_ab + thm2 +
  income_ab + thm2 +
  plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4)) &
  theme(legend.position = "bottom", 
        legend.box.margin = margin(t = -15, r = 0, b = 0, l = 0, unit = "pt"))
```
